A Critical Review of Sophistication and Model Selection in Algorithmic Statistics

### 2014-12-10 Facticity/Facticity/paper.tex
...

@@

@@

\section{Paul's sophistication (total, non-prefix) is \emph{not}  bounded}

@@

\begin{theorem}
There are infinitely many $x$ with 
\begin{enumerate}
  \item $S^\F_K(x) \geq K(x)/5$,\label{eq:item1}
  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
\end{enumerate}
\end{theorem}
\begin{proof}
From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic. For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic. Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings. Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$. By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$. This proves Item~\ref{eq:item1}.

For Items~\ref{eq:poezenvoer} and~\ref{eq:hondevoer}, suppose towards contradiction that either $C^\T_K(x)$ or $C^T_{\tau^\circ}(x)$ contains a candidate $f$ with $K(f)\le i/5$. Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\le\tau^\circ(x)\le K(x)$, so (a) above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) for large enough $i$. Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
\end{proof}
Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.

>>>>

\let\doendproof\endproof
\renewcommand\endproof{~\hfill\qed\doendproof}

@@

\begin{theorem}
There are infinitely many $x$ such that $S^\F_K(x) \geq K(x)/5$. \label{theorem:structure-function-is-not-bounded}
\end{theorem}
\begin{proof}
From \cite[Proposition~I.3 (b)]{gacs2001algorithmic} we know that for sufficiently large $i$ there are strings of length less than $i$ that are not $(i/4, i/4)$-stochastic. For each $i$, let $x_i$ be the smallest binary string that is not $(i/4, i/4)$-stochastic. Since every string is $(\alpha,\alpha)$-stochastic for sufficiently large $\alpha$, this sequence contains infinitely many distinct binary strings. Now pick any $i$ that is large enough that (a) $|x_i|<i$ and (b) $K(x_i)/5 < i/4$. By definition of $(\alpha,\beta)$-stochasticity, if a string is not $(i/4,i/4)$-stochastic, then for every finite set $S$ containing $x$, either (a) $\log|S|\ge K(x_i)+i/4$, in which case $S$ cannot be a candidate representation for $x$ in $C^\F_K$, or (b) $K(S)>i/4>K(x_i)/5$. This proves the theorem.
\end{proof}

@@

@@

We use this lemma to transport the result of Theorem~\ref{theorem:structure-function-is-not-bounded} to the model class of total functions:

\begin{theorem}
There are infinitely many $x$ with 
\begin{enumerate}
  \item $S^\T_K(x) \geq K(x)/6$, {and}\label{eq:poezenvoer}
  \item $S^\T_{\tau^\circ}(x) \geq K(x)/6$. \label{eq:hondevoer}
\end{enumerate}
\end{theorem}
\begin{proof}
We continue from the proof of Theorem~\ref{theorem:structure-function-is-not-bounded}. Suppose towards contradiction that either $C^\T_K(x_i)$ or $C^T_{\tau^\circ}(x_i)$ contains a candidate $f$ with $K(f)\le i/5$. Then by Lemma~\ref{lemma:total-to-sets}, there is a set $S$ with $\log|S|\le|d|\le\tau^\circ(x)\le K(x)$, so (a) in the proof of Theorem~\ref{theorem:structure-function-is-not-bounded} above cannot be the case, and $K(S)\le K(f)+K(|d|)+c\le i/5+K(|d|)+c$, which contradicts (b) in the same proof for large enough $i$. Therefore such candidates $f$ cannot exist and all candidates must satisfy $K(f)>i/5>K(x_i)/6$.
\end{proof}

Note that we favoured simplicity of the proof over the strength of the bounds: the constants can be improved.
